---
title: "ICUS vs CCUS analysis"
author: "Christelle Colin-Leitzinger"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: no
    theme: cerulean
    highlight: pygments
    df_print: paged
editor_options:
  chunk_output_type: console
---

<style type="text/css">
h1.title {
  font-size: 25px;
}
.figure {
   margin-top: 25px;
   margin-bottom: 25px;
}

table {
    margin-top: 25px;
    margin-bottom: 25px !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```


<style>
div.blue { background-color:#0099CC; border-radius: 5px; padding: 20px; font-size: 38px}
</style>
<div class = "blue">

<span style="color: white;">ICUS vs CCUS analysis</span>

</div>
<br>

```{r library, include=FALSE}
library(tidyverse)
library(viridis)
library(lubridate)
library(labelled)
library(gtsummary)
library(survival)
library(survminer)
theme_gtsummary_compact()
theme_set(theme_classic(base_size = 18))
```

```{r load}
path <- fs::path("","Volumes","Gillis_Research","Christelle Colin-Leitzinger", "CH and MMA")

variant <- 
  read_csv(paste0(here::here(), "/ICUS vs CCUS analysis/Variant_data_for_Padron_2023-08-28.csv"))

clinical <- 
  read_csv(paste0(here::here(), "/ICUS vs CCUS analysis/Clinical_data_for_Padron_2023-08-28.csv")) %>% 
  distinct()

MMA <- 
  readxl::read_xlsx(here::here(
    "ICUS vs CCUS analysis/Padron MMA and Succinate in human serum 12132023.xlsx"))
```

```{r}
med_data <- left_join(MMA %>% 
                    rename(SAMPLE_ID = s_sampleid), 
                  clinical, 
                  by = c("MDSEPID", "SAMPLE_ID")) %>% 
  janitor::clean_names() %>% 
  mutate(med_name = str_to_lower(med_name))

data <- left_join(MMA %>% 
                    rename(SAMPLE_ID = s_sampleid), 
                  clinical, 
                  by = c("MDSEPID", "SAMPLE_ID")) %>% 
  # Remove duplicate rows due to drugs
  distinct(MDSEPID, SAMPLE_ID, .keep_all = TRUE) %>% 
  left_join(., variant, 
                  by = c("MDSEPID", "SAMPLE_ID")) %>% 
  janitor::clean_names() %>% 
  rename(patient_id = mdsepid) %>% 
  # MMA----
  mutate(mma_um = mma_n_m / 1000) %>% 
  mutate(mma_cat = case_when(
    mma_um <= 0.4          ~ "≤0.4 ~ normal",
    mma_um > 0.4           ~ ">0.4 ~ high"
  ), mma_cat = factor(mma_cat, 
                      levels = c("≤0.4 ~ normal", ">0.4 ~ high"))
  ) %>% 
  # B12----
  mutate(b12_interpretation = case_when(
    baseline_b12_serum < 200             ~ "Low",
    baseline_b12_serum >= 200 &
      baseline_b12_serum <= 900          ~ "Normal",
    baseline_b12_serum > 900             ~ "High"
  ), b12_interpretation = factor(b12_interpretation, 
                                 levels = c("Low", "Normal", "High"))) %>% 
  # BMI----
  mutate(height_m = height_cm / 100,
         bmi = weight_kg / (height_m * height_m)) %>%
  mutate(bmi_cat = case_when(
    bmi < 18.5                  ~ "< 18.5",
    bmi >= 18.5 &
      bmi < 25                  ~ "18.5 - 25",
    bmi >= 25 &
      bmi < 30                  ~ "25 - 30",
    bmi >= 30                   ~ "≥ 30"
  )) %>%
  mutate(bmi_cat = factor(bmi_cat, levels = c("< 18.5", "18.5 - 25", "25 - 30", "≥ 30"))) %>% 
  # Now can Calculate ideal dody weight
  mutate(ideal_weight = case_when(
    sex == "Female" &
      height_m <= 1.524               ~ 45.5 + 2.3,
    sex == "Female" &
      height_m > 1.524                ~ 45.5 + (2.3 * (((height_m - 1.524) / 0.0254))),
    sex == "Male" &
      height_m <= 1.524               ~ 50 + 2.3,
    sex == "Male" &
      height_m > 1.524                ~ 50 + (2.3 * (((height_m - 1.524) / 0.0254)))
  )) %>% 
  mutate(adjusted_weight = ((weight_kg - ideal_weight) * 0.40) + ideal_weight
  ) %>%
  # CrCl----
  mutate(baseline_creatinine_serum = as.numeric(baseline_creatinine_serum),
         baseline_creatinine_serum = case_when(
           baseline_creatinine_serum < 0.7      ~ 0.7,
           TRUE                                   ~ baseline_creatinine_serum
         )) %>%
  # Calculate CrCl
  mutate(CrCl = case_when(
    sex == "Female"                   ~ ((
      ((140 - age) * weight_kg) /
        (72 * baseline_creatinine_serum)) * 
        0.85),
    
    sex == "Male"                     ~ (
      ((140 - age) * weight_kg) / 
        (72 * baseline_creatinine_serum)
      )
  ),
  CrCl = round(CrCl, 0)) %>% 
  
  mutate(adjusted_CrCl = case_when(
    sex == "Female"                   ~ ((
      ((140 - age) * adjusted_weight) /
        (72 * baseline_creatinine_serum)) * 
        0.85),
    
    sex == "Male"                     ~ (
      ((140 - age) * adjusted_weight) / 
        (72 * baseline_creatinine_serum)
      )
  ),
  adjusted_CrCl = round(adjusted_CrCl, 0)) %>% 

  mutate(calculated_CrCl = case_when(
    bmi >= 25                 ~ adjusted_CrCl,
    bmi < 25                  ~ CrCl
  )) %>% 
  mutate(CrCl_units = "mL/min/1.73m2") %>% 
  mutate(CrCl_interpretation = case_when(
    calculated_CrCl < 60                    ~ "Low",
    calculated_CrCl >= 60                   ~ "Normal"
  ), CrCl_interpretation = factor(CrCl_interpretation, 
                                  levels = c("Low", "Normal"))) %>% 
  
  # Age----
  mutate(age_cat = case_when(
    age < 50               ~ "< 50",
    age >= 50 & 
      age < 60             ~ "≥ 50 - 60",
    age >= 60 & 
      age < 70             ~ "≥ 60 - 70",
    age >= 70 & 
      age < 80             ~ "≥ 70 - 80",
    age >= 80 & 
      age < 90             ~ "≥ 80 - 90",
    age >= 90              ~ "≥ 90"
  )) %>% 
  mutate(age_cat2 = case_when(
    age < 50               ~ "< 50",
    age >= 50 & 
      age < 65             ~ "≥ 50 - 65",
    age >= 65              ~ "≥ 90"
  )) %>% 

  # OS----
  mutate(os_event = case_when(
    !is.na(death_dt)       ~ 1,
    is.na(death_dt)        ~ 0
  )) %>% 
  mutate_at(c("death_dt", "last_contact_dt", "start_dt"), ~ as.Date(., "%m/%d/%y")) %>% 
  mutate(os_date = coalesce(death_dt, last_contact_dt)) %>% 
  mutate(os_time = interval(start = start_dt, end = os_date) /
           duration(n = 1, units = "months"))
```

```{r Labeling}
var_label(data) <- list(age = "Patient age", 
                        sex = "Sex", 
                        bmi = "BMI", bmi_cat = "BMI categories", 
                        renal_disease = "Renal disease", 
                        mma_um = "MMA (µM)", mma_cat = "MMA (categories)",
                        succinate_ug_m_l = "Succinate (µg/ml)",
                        baseline_b12_serum = "Baseline B12 (pg/mL)",
                        b12_interpretation = "B12 (categories)", 
                        baseline_creatinine_serum = "Baseline creatinine (mg/dL)",
                        calculated_CrCl = "Creatinine clearance (mL/min/1.73m2)",
                        CrCl_interpretation = "Creatinine clearance (categories)",
                        age_of_the_samples_days = "Sample Storage Time (days)",
                        variant_confidence_level = "Variant Confidence Level")
```

# I. Table 1 - Clinical
```{r}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(age, sex,
         bmi, bmi_cat,
         mma_um, mma_cat,
         succinate_ug_m_l,
         baseline_b12_serum, 
         b12_interpretation, 
         baseline_creatinine_serum,
         renal_disease,
         calculated_CrCl,
         CrCl_interpretation,
         group) %>% 
  tbl_summary(by = group) %>% 
  bold_labels() %>% add_overall() %>% 
  add_p() %>% bold_p(t = 0.05)

med_data %>% 
  select(med_name, med_reason,
         group) %>% 
  mutate(across(starts_with("med_"), ~ str_to_sentence(.))) %>% 
  tbl_summary(by = group, 
              sort = everything() ~ "frequency") %>% 
  bold_labels() %>% add_overall() %>% 
  add_p() %>% bold_p(t = 0.05)
```

## Creatinine overview
```{r}
data %>% 
  filter(!is.na(baseline_creatinine_serum)) %>% 
  ggplot(aes(x= baseline_creatinine_serum))+
  geom_histogram(binwidth = 0.1)+
  labs(x = "Baseline creatinine (mg/dL)", 
       subtitle = "creatinine was cleaned as :
       values < 0.7 were set to be equal to 0.7")+
  scale_x_continuous(limits = c(0.7, 6),
                     breaks = c(0.7, 2, 4, 6))
data %>% 
  filter(!is.na(calculated_CrCl)) %>% 
  ggplot(aes(x= calculated_CrCl))+
  geom_histogram(binwidth = 1)+
  labs(x = "Creatinine clearance (mL/min/1.73m2)")
```

## B12 overview
```{r}
data %>% 
  filter(!is.na(baseline_b12_serum)) %>% 
  ggplot(aes(x= baseline_b12_serum))+
  geom_histogram(binwidth = 100)+
  labs(x = "Baseline B12 (pg/mL)")
```

# II. Table 2 - Variant
## ALL variants of all patients
1 patient is categorized as ICUS but has a known mutation.
```{r}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(variant_confidence_level,
         group) %>% 
  tbl_summary(by = group) %>% 
  bold_labels() %>% add_overall() %>% 
  add_p() %>% bold_p(t = 0.05)

data %>% 
  filter(variant_confidence_level == "Level 1 (Manually Curated)" &
           group == "ICUS_GROUP") %>% 
  select(patient_id, group,
         gene : variant_confidence_level, 
         everything())
```

## Mutations count summarized by gene
```{r}
# data %>% 
#   filter(!is.na(gene)) %>% 
#   # Count has patient - remove duplicate gene per patient
#   distinct(patient_id, gene) %>% 
#   select(gene) %>% 
#   tbl_summary(sort = everything() ~ "frequency") %>% 
#   bold_labels()

library(fastDummies)
dum <- data %>% 
  select(patient_id, gene) %>% 
  filter(!is.na(gene)) %>% 
  distinct(patient_id, gene)
dum %>% 
  dummy_cols(select_columns = "gene") %>% 
  select(-gene) %>% 
  mutate(across(starts_with("gene"), ~ na_if(., 0))) %>% 
  group_by(patient_id) %>% 
  fill(everything(), .direction = "updown") %>% 
  ungroup() %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  mutate(across(starts_with("gene"), .fns = ~ replace_na(., 0))) %>% 
  select(-patient_id) %>% 
  `colnames<-`(str_remove(colnames(.), "gene_")) %>% 
  tbl_summary(sort = everything() ~ "frequency") %>% 
  bold_labels()
```


# III. Explore previous analysis
## 1. CCUS vs. ICUS
<div class = "row">
<div class = "col-md-6">
MMA
```{r}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = group, y = mma_um))+
  geom_violin()+
  geom_jitter(color="grey")+
  geom_boxplot(width=0.1, outlier.shape = NA)+
  labs(x = NULL, y = "MMA (µM)")

data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(group, mma_um, mma_cat) %>% 
  tbl_summary(by = group) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>

<div class = "col-md-6">
Succinate
```{r}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = group, y = succinate_ug_m_l))+
  geom_violin()+
  geom_jitter(color="grey")+
  geom_boxplot(width=0.1, outlier.shape = NA)+
  labs(x = NULL, y = "Succinate (ug/ml)")

data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(group, succinate_ug_m_l) %>% 
  tbl_summary(by = group) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>
</div>


## 2. Age
<div class = "row">
<div class = "col-md-6">
MMA
```{r}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = age, y = mma_um))+
  geom_point(aes(color = age_cat))+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 1)+ 
  stat_regline_equation(label.y = 0.75)+
  labs(x = "Age (continuous)", y = "MMA (µM)")

data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(age_cat, mma_um, mma_cat) %>% 
  tbl_summary(by = age_cat) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>

<div class = "col-md-6">
Succinate
```{r}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = age, y = succinate_ug_m_l))+
  geom_point(aes(color = age_cat))+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 30)+ 
  stat_regline_equation(label.y = 20)+
  labs(x = "Age (continuous)", y = "Succinate (ug/ml)")

data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(age_cat, succinate_ug_m_l) %>% 
  tbl_summary(by = age_cat) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>
</div>

## 3. Collection date
<div class = "row">
<div class = "col-md-6">
MMA
```{r}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = age_of_the_samples_days, y = mma_um))+
  geom_point(color = "blue")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 1)+ 
  stat_regline_equation(label.y = 0.75)+
  labs(x = "Sample Storage Time (days)", y = "MMA (µM)")
```
</div>

<div class = "col-md-6">
Succinate
```{r}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = age_of_the_samples_days, y = succinate_ug_m_l))+
  geom_point(color = "blue")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 30)+ 
  stat_regline_equation(label.y = 20)+
  labs(x = "Sample Storage Time (days)", y = "Succinate (ug/ml)")
```
</div>
</div>

## 4. Sex
<div class = "row">
<div class = "col-md-6">
MMA
```{r}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = sex, y = mma_um))+
  geom_violin()+
  geom_jitter(color="grey")+
  geom_boxplot(width=0.1, outlier.shape = NA)+
  labs(x = NULL, y = "MMA (µM)")

data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(sex, mma_um, mma_cat) %>% 
  tbl_summary(by = sex) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>

<div class = "col-md-6">
Succinate
```{r}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = sex, y = succinate_ug_m_l))+
  geom_violin()+
  geom_jitter(color="grey")+
  geom_boxplot(width=0.1, outlier.shape = NA)+
  labs(x = NULL, y = "Succinate (ug/ml)")

data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(sex, succinate_ug_m_l) %>% 
  tbl_summary(by = sex) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>
</div>

# IV. New analysis
## 1. Survival in overall population
```{r}
surv_data <- data %>% 
  distinct(patient_id, .keep_all = TRUE)
  
ggsurvplot(survfit(Surv(os_time, os_event) ~ group, 
                   data=surv_data),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1))

tbl1 <- surv_data %>% 
  select(group, 
         mma_um,
         age, sex, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with MMA in overall data - continuous MMA"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + mma_um + age + sex + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))

tbl1 <- surv_data %>% 
  select(group, 
         mma_cat,
         age, sex, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with MMA in overall data - categorical MMA"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + mma_cat + age + sex + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))

tbl1 <- surv_data %>% 
  select(group, 
         succinate_ug_m_l,
         age, sex, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with succinate in overall data"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + succinate_ug_m_l + age + sex + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

## 2. Stratified by age
<div class = "row">
<div class = "col-md-6">
MMA
```{r, fig.height=8}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = age, y = mma_um))+
  geom_point(aes(color = age_cat2))+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 1)+ 
  stat_regline_equation(label.y = 0.75)+
  labs(x = "Age", y = "MMA (µM)")+
  facet_wrap(. ~ age_cat2, ncol = 1)

data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(age_cat2, mma_um, mma_cat) %>% 
  tbl_summary(by = age_cat2) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>

<div class = "col-md-6">
Succinate
```{r, fig.height=8}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = age, y = succinate_ug_m_l))+
  geom_point(aes(color = age_cat2))+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 30)+ 
  stat_regline_equation(label.y = 20)+
  labs(x = "Age", y = "Succinate (ug/ml)")+
  facet_wrap(. ~ age_cat2, ncol = 1)

data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(age_cat2, succinate_ug_m_l) %>% 
  tbl_summary(by = age_cat2) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>
</div>


<div class = "row">
<div class = "col-md-6">
MMA
```{r}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(age > 50) %>% 
  ggplot(aes(x = age, y = mma_um))+
  geom_point(color = "purple")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 1)+ 
  stat_regline_equation(label.y = 0.75)+
  labs(x = "Age", y = "MMA (µM)", 
       subtitle = "Age > 50")

data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(age > 60) %>% 
  ggplot(aes(x = age, y = mma_um))+
  geom_point(color = "purple")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 1)+ 
  stat_regline_equation(label.y = 0.75)+
  labs(x = "Age", y = "MMA (µM)", 
       subtitle = "Age > 60")

data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(age > 65) %>% 
  ggplot(aes(x = age, y = mma_um))+
  geom_point(color = "purple")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 1)+ 
  stat_regline_equation(label.y = 0.75)+
  labs(x = "Age", y = "MMA (µM)", 
       subtitle = "Age > 65")
```
</div>

<div class = "col-md-6">
Succinate
```{r}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(age > 50) %>% 
  ggplot(aes(x = age, y = succinate_ug_m_l))+
  geom_point(color = "orange")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 30)+ 
  stat_regline_equation(label.y = 20)+
  labs(x = "Age", y = "Succinate (ug/ml)")

data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(age > 60) %>% 
  ggplot(aes(x = age, y = succinate_ug_m_l))+
  geom_point(color = "orange")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 30)+ 
  stat_regline_equation(label.y = 20)+
  labs(x = "Age", y = "Succinate (ug/ml)")

data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(age > 65) %>% 
  ggplot(aes(x = age, y = succinate_ug_m_l))+
  geom_point(color = "orange")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 30)+ 
  stat_regline_equation(label.y = 20)+
  labs(x = "Age", y = "Succinate (ug/ml)")
```
</div>
</div>

## 3. Stratified by Sex
### Male
```{r}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(sex == "Male") %>% 
  select(mma_um, mma_cat,
         succinate_ug_m_l,
         group) %>% 
  tbl_summary(by = group) %>% 
  bold_labels() %>% add_overall() %>% 
  add_p() %>% bold_p(t = 0.05)
```

```{r}
surv_data <- data %>% 
  filter(sex == "Male") %>% 
  distinct(patient_id, .keep_all = TRUE)
  
ggsurvplot(survfit(Surv(os_time, os_event) ~ group, 
                   data=surv_data),
           title = "OS Analysis in Male",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1))

tbl1 <- surv_data %>% 
  select(group, 
         mma_um,
         age, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with MMA in male data - continuous MMA"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + mma_um + age + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))

tbl1 <- surv_data %>% 
  select(group, 
         mma_cat,
         age, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with MMA in male data - categorical MMA"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + mma_cat + age + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))

tbl1 <- surv_data %>% 
  select(group, 
         succinate_ug_m_l,
         age, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with succinate in male data"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + succinate_ug_m_l + age + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```


### Female
```{r}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(sex == "Female") %>% 
  select(mma_um, mma_cat,
         succinate_ug_m_l,
         group) %>% 
  tbl_summary(by = group) %>% 
  bold_labels() %>% add_overall() %>% 
  add_p() %>% bold_p(t = 0.05)
```

```{r}
surv_data <- data %>% 
  filter(sex == "Female") %>% 
  distinct(patient_id, .keep_all = TRUE)
  
ggsurvplot(survfit(Surv(os_time, os_event) ~ group, 
                   data=surv_data),
           title = "OS Analysis in Female",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1))

tbl1 <- surv_data %>% 
  select(group, 
         mma_um,
         age, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with MMA in female data - continuous MMA"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + mma_um + age + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))

tbl1 <- surv_data %>% 
  select(group, 
         mma_cat,
         age, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with MMA in male data - categorical MMA"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + mma_cat + age + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))

tbl1 <- surv_data %>% 
  select(group, 
         succinate_ug_m_l,
         age, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with succinate in female data"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + succinate_ug_m_l + age + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

## 4. MMA & B12
```{r}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = baseline_b12_serum, y = mma_um))+
  geom_point(color = "blue")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 1)+ 
  stat_regline_equation(label.y = 0.75)+
  labs(x = "Baseline B12 (pg/mL)", y = "MMA (µM)")

data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(!is.na(b12_interpretation)) %>% 
  ggplot(aes(x = b12_interpretation, y = mma_um))+
  geom_boxplot()+
  geom_jitter(color="grey")

data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(b12_interpretation, mma_um) %>% 
  tbl_summary(by = b12_interpretation) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```

## 5. MMA & kidney function (renal disease)
<div class = "row">
<div class = "col-md-6">
```{r}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(!is.na(renal_disease)) %>% 
  ggplot(aes(x = renal_disease, y = mma_um))+
  geom_violin()+
  geom_jitter(color="grey")+
  geom_boxplot(width=0.1, outlier.shape = NA)+
  labs(x = "Renal disease", y = "MMA (µM)")

data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(renal_disease, mma_um) %>% 
  tbl_summary(by = renal_disease) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>

<div class = "col-md-6">
```{r}
data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(!is.na(CrCl_interpretation)) %>% 
  ggplot(aes(x = CrCl_interpretation, y = mma_um))+
  geom_violin()+
  geom_jitter(color="grey")+
  geom_boxplot(width=0.1, outlier.shape = NA)+
  labs(x = "Creatinine clearance", y = "MMA (µM)")

data %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(CrCl_interpretation, mma_um) %>% 
  tbl_summary(by = CrCl_interpretation) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>
</div>


# V. CCUS vs ICUS MMA when exclude patients on B12 supplementation
```{r}
b12_excluded <- med_data %>% 
  rename(patient_id = mdsepid) %>% 
  mutate(drug_ever = case_when(
    !is.na(med_name)            ~ "Yes",
    !is.na(med_start_dt)        ~ "Yes"
  )) %>% 
  group_by(patient_id) %>% 
  fill(drug_ever, .direction = "updown") %>% 
  ungroup() %>% 
  filter(is.na(drug_ever))

b12_excluded <- data %>% 
  filter(str_detect(patient_id, paste0(b12_excluded$patient_id, collapse = "|")))
```

## I. Table 1 - Clinical
```{r}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(age, sex,
         bmi, bmi_cat,
         mma_um, mma_cat,
         succinate_ug_m_l,
         baseline_b12_serum, 
         b12_interpretation, 
         baseline_creatinine_serum,
         renal_disease,
         calculated_CrCl,
         CrCl_interpretation,
         group) %>% 
  tbl_summary(by = group) %>% 
  bold_labels() %>% add_overall() %>% 
  add_p() %>% bold_p(t = 0.05)
```

## II. Table 2 - Variant
### ALL variants of all patients
The patient previously categorized as ICUS was excluded.
```{r}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(variant_confidence_level,
         group) %>% 
  tbl_summary(by = group) %>% 
  bold_labels() %>% add_overall() %>% 
  add_p() %>% bold_p(t = 0.05)

# b12_excluded %>% 
#   filter(variant_confidence_level == "Level 1 (Manually Curated)" &
#            group == "ICUS_GROUP") %>% 
#   select(patient_id, group,
#          gene : variant_confidence_level, 
#          everything())
```

### Mutations count summarized by gene
```{r}
library(fastDummies)
dum <- b12_excluded %>% 
  select(patient_id, gene) %>% 
  filter(!is.na(gene)) %>% 
  distinct(patient_id, gene)
dum %>% 
  dummy_cols(select_columns = "gene") %>% 
  select(-gene) %>% 
  mutate(across(starts_with("gene"), ~ na_if(., 0))) %>% 
  group_by(patient_id) %>% 
  fill(everything(), .direction = "updown") %>% 
  ungroup() %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  mutate(across(starts_with("gene"), .fns = ~ replace_na(., 0))) %>% 
  select(-patient_id) %>% 
  `colnames<-`(str_remove(colnames(.), "gene_")) %>% 
  tbl_summary(sort = everything() ~ "frequency") %>% 
  bold_labels()
```

## III. Explore previous analysis
### 1. CCUS vs. ICUS
<div class = "row">
<div class = "col-md-6">
MMA
```{r}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = group, y = mma_um))+
  geom_violin()+
  geom_jitter(color="grey")+
  geom_boxplot(width=0.1, outlier.shape = NA)+
  labs(x = NULL, y = "MMA (µM)")

b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(group, mma_um, mma_cat) %>% 
  tbl_summary(by = group) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>

<div class = "col-md-6">
Succinate
```{r}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = group, y = succinate_ug_m_l))+
  geom_violin()+
  geom_jitter(color="grey")+
  geom_boxplot(width=0.1, outlier.shape = NA)+
  labs(x = NULL, y = "Succinate (ug/ml)")

b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(group, succinate_ug_m_l) %>% 
  tbl_summary(by = group) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>
</div>


### 2. Age
<div class = "row">
<div class = "col-md-6">
MMA
```{r}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = age, y = mma_um))+
  geom_point(aes(color = age_cat))+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 1)+ 
  stat_regline_equation(label.y = 0.75)+
  labs(x = "Age (continuous)", y = "MMA (µM)")

b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(age_cat, mma_um, mma_cat) %>% 
  tbl_summary(by = age_cat) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>

<div class = "col-md-6">
Succinate
```{r}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = age, y = succinate_ug_m_l))+
  geom_point(aes(color = age_cat))+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 30)+ 
  stat_regline_equation(label.y = 20)+
  labs(x = "Age (continuous)", y = "Succinate (ug/ml)")

b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(age_cat, succinate_ug_m_l) %>% 
  tbl_summary(by = age_cat) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>
</div>

### 3. Collection date
<div class = "row">
<div class = "col-md-6">
MMA
```{r}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = age_of_the_samples_days, y = mma_um))+
  geom_point(color = "blue")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 1)+ 
  stat_regline_equation(label.y = 0.75)+
  labs(x = "Sample Storage Time (days)", y = "MMA (µM)")
```
</div>

<div class = "col-md-6">
Succinate
```{r}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = age_of_the_samples_days, y = succinate_ug_m_l))+
  geom_point(color = "blue")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 30)+ 
  stat_regline_equation(label.y = 20)+
  labs(x = "Sample Storage Time (days)", y = "Succinate (ug/ml)")
```
</div>
</div>

### 4. Sex
<div class = "row">
<div class = "col-md-6">
MMA
```{r}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = sex, y = mma_um))+
  geom_violin()+
  geom_jitter(color="grey")+
  geom_boxplot(width=0.1, outlier.shape = NA)+
  labs(x = NULL, y = "MMA (µM)")

b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(sex, mma_um, mma_cat) %>% 
  tbl_summary(by = sex) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>

<div class = "col-md-6">
Succinate
```{r}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = sex, y = succinate_ug_m_l))+
  geom_violin()+
  geom_jitter(color="grey")+
  geom_boxplot(width=0.1, outlier.shape = NA)+
  labs(x = NULL, y = "Succinate (ug/ml)")

b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(sex, succinate_ug_m_l) %>% 
  tbl_summary(by = sex) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>
</div>

## IV. New analysis
### 1. Survival in overall population
```{r}
surv_data <- b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE)
  
ggsurvplot(survfit(Surv(os_time, os_event) ~ group, 
                   data=surv_data),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1))

tbl1 <- surv_data %>% 
  select(group, 
         mma_um,
         age, sex, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with MMA in overall data - continuous MMA"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + mma_um + age + sex + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))

tbl1 <- surv_data %>% 
  select(group, 
         mma_cat,
         age, sex, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with MMA in overall data - categorical MMA"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + mma_cat + age + sex + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))

tbl1 <- surv_data %>% 
  select(group, 
         succinate_ug_m_l,
         age, sex, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with succinate in overall data"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + succinate_ug_m_l + age + sex + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

### 2. Stratified by age
<div class = "row">
<div class = "col-md-6">
MMA
```{r, fig.height=8}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = age, y = mma_um))+
  geom_point(aes(color = age_cat2))+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 1)+ 
  stat_regline_equation(label.y = 0.75)+
  labs(x = "Age", y = "MMA (µM)")+
  facet_wrap(. ~ age_cat2, ncol = 1)

b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(age_cat2, mma_um, mma_cat) %>% 
  tbl_summary(by = age_cat2) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>

<div class = "col-md-6">
Succinate
```{r, fig.height=8}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = age, y = succinate_ug_m_l))+
  geom_point(aes(color = age_cat2))+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 30)+ 
  stat_regline_equation(label.y = 20)+
  labs(x = "Age", y = "Succinate (ug/ml)")+
  facet_wrap(. ~ age_cat2, ncol = 1)

b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(age_cat2, succinate_ug_m_l) %>% 
  tbl_summary(by = age_cat2) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>
</div>


<div class = "row">
<div class = "col-md-6">
MMA
```{r}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(age > 50) %>% 
  ggplot(aes(x = age, y = mma_um))+
  geom_point(color = "purple")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 1)+ 
  stat_regline_equation(label.y = 0.75)+
  labs(x = "Age", y = "MMA (µM)")

b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(age > 60) %>% 
  ggplot(aes(x = age, y = mma_um))+
  geom_point(color = "purple")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 1)+ 
  stat_regline_equation(label.y = 0.75)+
  labs(x = "Age", y = "MMA (µM)")

b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(age > 65) %>% 
  ggplot(aes(x = age, y = mma_um))+
  geom_point(color = "purple")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 1)+ 
  stat_regline_equation(label.y = 0.75)+
  labs(x = "Age", y = "MMA (µM)")
```
</div>

<div class = "col-md-6">
Succinate
```{r}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(age > 50) %>% 
  ggplot(aes(x = age, y = succinate_ug_m_l))+
  geom_point(color = "orange")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 30)+ 
  stat_regline_equation(label.y = 20)+
  labs(x = "Age", y = "Succinate (ug/ml)")

b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(age > 60) %>% 
  ggplot(aes(x = age, y = succinate_ug_m_l))+
  geom_point(color = "orange")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 30)+ 
  stat_regline_equation(label.y = 20)+
  labs(x = "Age", y = "Succinate (ug/ml)")

b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(age > 65) %>% 
  ggplot(aes(x = age, y = succinate_ug_m_l))+
  geom_point(color = "orange")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 30)+ 
  stat_regline_equation(label.y = 20)+
  labs(x = "Age", y = "Succinate (ug/ml)")
```
</div>
</div>

### 3. Stratified by Sex
#### Male
```{r}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(sex == "Male") %>% 
  select(mma_um, mma_cat,
         succinate_ug_m_l,
         group) %>% 
  tbl_summary(by = group) %>% 
  bold_labels() %>% add_overall() %>% 
  add_p() %>% bold_p(t = 0.05)
```

```{r}
surv_data <- b12_excluded %>% 
  filter(sex == "Male") %>% 
  distinct(patient_id, .keep_all = TRUE)
  
ggsurvplot(survfit(Surv(os_time, os_event) ~ group, 
                   data=surv_data),
           title = "OS Analysis in Male",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1))

tbl1 <- surv_data %>% 
  select(group, 
         mma_um,
         age, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with MMA in male data - continuous MMA"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + mma_um + age + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))

tbl1 <- surv_data %>% 
  select(group, 
         mma_cat,
         age, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with MMA in male data - categorical MMA"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + mma_cat + age + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))

tbl1 <- surv_data %>% 
  select(group, 
         succinate_ug_m_l,
         age, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with succinate in male data"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + succinate_ug_m_l + age + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```


#### Female
```{r}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(sex == "Female") %>% 
  select(mma_um, mma_cat,
         succinate_ug_m_l,
         group) %>% 
  tbl_summary(by = group) %>% 
  bold_labels() %>% add_overall() %>% 
  add_p() %>% bold_p(t = 0.05)
```

```{r}
surv_data <- b12_excluded %>% 
  filter(sex == "Female") %>% 
  distinct(patient_id, .keep_all = TRUE)
  
ggsurvplot(survfit(Surv(os_time, os_event) ~ group, 
                   data=surv_data),
           title = "OS Analysis in Female",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1))

tbl1 <- surv_data %>% 
  select(group, 
         mma_um,
         age, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with MMA in female data - continuous MMA"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + mma_um + age + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))

tbl1 <- surv_data %>% 
  select(group, 
         mma_cat,
         age, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with MMA in male data - categorical MMA"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + mma_cat + age + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)
tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))

tbl1 <- surv_data %>% 
  select(group, 
         succinate_ug_m_l,
         age, bmi, 
         CrCl_interpretation,
         baseline_b12_serum) %>% 
  tbl_uvregression(method = survival::coxph, 
                   y = (Surv(time = surv_data$os_time, 
                             event = surv_data$os_event)),
                   exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels() %>% 
  modify_spanning_header(everything() ~ paste("Association with succinate in female data"))
tbl2 <- coxph(Surv(time = surv_data$os_time, 
             event = surv_data$os_event) ~ 
               group + mma_um + succinate_ug_m_l + age + bmi + CrCl_interpretation + baseline_b12_serum, 
             data =  surv_data) %>%
  tbl_regression(exponentiate = TRUE) %>% bold_p(t = .05)

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

### 4. MMA & B12
```{r}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = baseline_b12_serum, y = mma_um))+
  geom_point(color = "blue")+
  geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x) +
  stat_cor(label.y = 1)+ 
  stat_regline_equation(label.y = 0.75)+
  labs(x = "Baseline B12 (pg/mL)", y = "MMA (µM)")

b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(!is.na(b12_interpretation)) %>% 
  ggplot(aes(x = b12_interpretation, y = mma_um))+
  geom_boxplot()+
  geom_jitter(color="grey")

b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(b12_interpretation, mma_um) %>% 
  tbl_summary(by = b12_interpretation) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```

### 5. MMA & kidney function (renal disease)
<div class = "row">
<div class = "col-md-6">
```{r}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(!is.na(renal_disease)) %>% 
  ggplot(aes(x = renal_disease, y = mma_um))+
  geom_violin()+
  geom_jitter(color="grey")+
  geom_boxplot(width=0.1, outlier.shape = NA)

b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(renal_disease, mma_um) %>% 
  tbl_summary(by = renal_disease) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>

<div class = "col-md-6">
```{r}
b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  filter(!is.na(CrCl_interpretation)) %>% 
  ggplot(aes(x = CrCl_interpretation, y = mma_um))+
  geom_violin()+
  geom_jitter(color="grey")+
  geom_boxplot(width=0.1, outlier.shape = NA)

b12_excluded %>% 
  distinct(patient_id, .keep_all = TRUE) %>% 
  select(CrCl_interpretation, mma_um) %>% 
  tbl_summary(by = CrCl_interpretation) %>% 
  bold_labels() %>% 
  add_p() %>% bold_p(t = 0.05)
```
</div>
</div>



